Make any stupid plot you want

### 2021-04-16 draft/sections/math.tex
finished out math section

\subsubsection{Variables in Fiber Space \dfiber}
\label{sec:math:data:fiber}
To formalize the structure of the data components, we use notation introduced by Spivak \cite{spivakSIMPLICIALDATABASES} that binds the components of the fiber to variable names. This allows us to describe the components in a schema like way. Spivak constructs a set \ftotal\ that is the disjoint union of all possible objects of types $\{\ftype_0, \ldots, \ftype_m\} \in \ftypes$, where \ftypes\ are the data types of the variables in the dataset. He then defines the single variable set \fttype\ 
\begin{equation}
    \label{eq:data_types}
    \begin{tikzcd}
        \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] & \ftotal \arrow[d, "\pi"] \\
        \fnames \arrow[r, "\fsection"']                          & \ftypes       
    \end{tikzcd}
\end{equation}
which is \ftotal\ restricted to objects of type \ftype\ bound to variable name \fname. The \fttype\ lookup is by name to specify that every component is distinct, since multiple components can have the same type \ftype. Given \fsection, the fiber for a one variable dataset is
\begin{equation}
    \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
\end{equation}
where \fsection\ is the schema that binds a variable name \fname\ to its datatype \ftype. A dataset with multiple components has a fiber that is the cartesian cross product of $\ftotal_{\fsection}$ applied to all the columns:
\begin{equation}
F = \ftotal_{\fsection(\fname_{1})}\times \ldots \ftotal_{\fsection(\fname_{i})} \ldots\times \ftotal_{\fsection(\fname_{n})}
\end{equation}
which is equivalent to 
\begin{equation}
    \dfiber= \dfiber_{0} \times \ldots \times \dfiber_{i}\times\ldots\times \dfiber_{n}
\end{equation}
which allows us to decouple \dfiber\ into components $\dfiber_i$.

\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/fiber.png}
    \caption{The fiber space is the set of all combinations of all theoretically possible values of the components.  The 2D fiber \(\dfiber=\reals\times\reals\) encodes the properties of \textit{time} and \textit{temperature} components.  One dimension of the fiber encodes the range of possible values for the time component of the dataset, which is a subset of the \reals, while the other dimension encodes the range of possible values  \reals\ for the temperature component. This means the fiber is the set of points \textit{(temperature, time)}that are all the combinations of \textit{temperatature} $\times$ \textit{time}. The 3D fiber encodes points at all possible combinations of \textit{temperature}, \textit{latitude}, and \textit{longitude}.}
    \label{fig:data:fiber:example}
\end{figure}

@@

As illustrated in figure~\ref{fig:data_fiber_example}, Spivak's framework provides a consistent way to describe potentially complex components of the input data.

\subsubsection{Measurement Scales: Monoid Actions}
\label{sec:data_monoid}
Implementing expressive visual encodings requires formally describing the structure on the components of the fiber, which we define by the actions of a monoid on the component. In doing so, we specify the properties of the component that must be preserved in a graphic representation. A monoid \cite{Monoid2021} $\monoid$ is a set with a binary operation $\ast:\monoid \times \monoid\rightarrow \monoid$ that satisfies the axioms:
\begin{align*}
    \textbf{associativity}\; & \text{for all} a, b, c \in \monoid \, (a \bullet b) \bullet c = a \bullet (b \bullet c)\\
    \textbf{identity}\; & \text{for all } a\in \monoid,\,  e\bullet a = a 
\end{align*} 
As defined on a component of \dfiber, a left monoid action \cite{SemigroupAction2021,nlab:action} of $\monoid_i$ is a set $\dfiber_i$ with an action $\bullet: \monoid\times \dfiber_i \rightarrow \dfiber_i$ with the properties:
\begin{align*}
    \textbf{associativity}\;& \text{for all } f,g \in \monoid_i \text{ and } x\in \dfiber_i,\, f\bullet(g\bullet x) = (f\ast g) \bullet x\\
    \textbf{identity}\;& \text{for all } x\in \dfiber_i, e\in \monoid_i,\,  e\bullet x = x 
\end{align*}
The identity and associativity properties of the action denote that the action is a monoid homomorphism, which means that the group operation is preserved on both sides of the action\cite{weissteinGroupHomomorphism}.

\begin{figure}[H]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/monoid_total.png}
        \caption{}
        \label{fig:monoid:total}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=1\textwidth]{figures/math/monoid_partial.png}
        \caption{}
        \label{fig:monoid:partial}
    \end{subfigure}
    \caption{The action $\bullet$ in \autoref{fig:monoid:total} is the arrows from the partial order diagram of weather states on the left to the diagram of weather states on the right. Since the action maps the weather states to themselves, the ordering defined by the monoid $\ast$ is preserved on both sides of the action. The action in \autoref{fig:monoid:partial} is monoid homomorphism because the ordering of the weather states is the same as the ordering of the elements they are mapped to. Given $sun \geq clouds\geq rain\, lightining$ on the right, the action  $sun\, clouds \rightarrow clouds$, and $rain\, lightining \rightarrow rain$ is structure preserving because on the left $cloud \geq rain$ so the relative ordering of elements is the same as the elements they are mapped to.}
    \label{fig:monoid}
\end{figure}

One example of monoids are partial orderings on a set, such as seen in \label{fig:monoid}. Each hasse diagram of the set of weather states describes an ordering on the set; the arrow goes from the lesser value to the greater one. For example, $storm \leq rain$. In \autoref{fig:monoid:total}, the action $\bullet$ maps the elements of a set of weather states into itself by mapping them into other elements of the weather states. The action in  \autoref{fig:monoid:total}, represented as the arrows between the hasse diagrams of the weather states, maps the weather states to themselves; therefore the ordering of the weather states is identical on both sides of the action and it is therefore homomorphic. The action $\bullet$ in \autoref{fig:monoid:partial} is a monotone map\cite{fongInvitationAppliedCategory2019}, meaning that for  $a,b \in \dfiber_{i}$
\begin{equation}
    if\, a \leq b \; then \; \bullet(a) \leq \bullet(b)
\end{equation}
the structure the action must preserve is the relative, rather than exact, ordering. Since groups are monoids with invertible operations, this definition of structure is also broad enough to include the Steven's measurment scales\cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale}. Monoids are also commonly found in functional programming because a property of monoids is that the components can be composed into complex transformation pipelines \cite{yorgeyMonoidsThemeVariations}.

As with the fiber \dfiber\, the total monoid space \monoid\ is the cartesian product
\begin{equation}
\monoid= \monoid_{0} \times \ldots \times \monoid_{i}\times \ldots \times\ldots \monoid_{n}
\end{equation}
of each monoid $\monoid_{i}$ on $\dfiber_{i}$.  The monoid is also added to the specification of the fiber $(\fname_i,\, \ftype_i,\, \fttype\, \monoid_i)$

\subsubsection{Continuity of the Data $K$} 
\label{sec:data_base}
The base space \dbase\ is way to express how the records in \dtotal\ are connected to each other, for example if they are discrete points or if they lie in a 2D continous surface. Connectivity type is assumed in the choice of visualization, for example a line plot implies 1D continuous data, but an explicit representation allows for verifying that the topology of the graphic representation is equivalent to the topology of the data.

\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/k_different_types.png}
    \caption{The topological base space \dbase\ encodes the continuity of the data space, for example if the data is discrete points or lies on a plane or a sphere}
    \label{fig:base_space_types}
\end{figure}
As illustrated in figure~\ref{fig:base_space_types}, \dbase\ is akin to an indexing space into \dtotal\ that describes the structure of \dtotal.  \dbase\ can have any number of dimensions and can be continuous or discrete.

\begin{figure}[H]
    \includegraphics[width=\linewidth]{figures/math/k_qspace.png}
    \caption{The base space \dtotal\ is divided into fiber segments \dfiber. The base space \dbase\ acts as an index into the records in the fibers.}
    \label{fig:base_space_div}
\end{figure}

Formally \dbase\ is the quotient space \cite{QuotientSpaceTopology2020} of \dtotal\, meaning it is the finest space\cite{aurouxMath131Introduction} such that every $\dbasepoint \in \dbase$ has a corresponding fiber $\dfiber_k$\cite{QuotientSpaceTopology2020}. In figure~\ref{fig:base_space_div}, \dtotal\ is a rectangle divided by vertical fibers \dfiber, so the minimal \dbase\ for which there is always a mapping $\pi: \dtotal\rightarrow \dbase$ is the closed interval $\left[0,1\right]$. 
As with fibers and monoids, we can decompose the total space into components $\pi:\dtotal_i\rightarrow \dbase$ where
\begin{equation}
    \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
\end{equation}
which is a decomposition of \dfiber. The \dbase\ remains the same because the connectivity of records does not change just because there are fewer elements in each record.
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/base.png}
    \caption{These two datasets have the same (time, temperature) fiber, but different continuities. The dataset on the left consists of discrete records, while the records in the dataset on the right sampled from a continuous space.} 
    \label{fig:base_example}
\end{figure}

The datasets in figure~\ref{fig:base_example} have the same fiber of (temperature, time). The dot represents a discrete base space \dbase, meaning that every dataset encoded in the fiber bundle has discrete continuity. The line is a representation of a 1D continuity, meaning that every dataset in the fiber bundle is 1D continous. By encoding this continuity in the model as \dbase\, the data model now explicitly carries information about its structure such that the implicit assumptions of the visualization algorithms are now explicit. The explicit topology is a concise way of distinguishing visualizations that appear identical, for example heatmaps and images.

\subsubsection{Data \dsection}
\label{sec:data_section}
While the projection function $\pi:\dtotal \rightarrow\dbase$ ties together the base space \dbase\ with the fiber \dfiber, a section $\dsection: \dbase\rightarrow \dtotal$ encodes a dataset. A section function takes as input location $\dbasepoint \in \dbase$ and returns a record $\delement \in \dtotal$. For example, in the special case of a table \cite{spivakSIMPLICIALDATABASES}, \dbase\ is a set of row ids, \dfiber\ is the columns, and the section \dsection\ returns the record \delement\ at a given key in \dbase. For any fiber bundle, there exists a map
\begin{equation}
    \begin{tikzcd}
        \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                          & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
 such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$. The set of all global sections is denoted as $\Gamma(\dtotal)$. Assuming a trivial fiber bundle $\dtotal = \dbase \times \dfiber$, the section is 
\begin{equation}
    \label{eq:section_return}
    \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
\end{equation}
where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber. This formulation of the section also holds on locally trivial sections of a non-trivial fiber bundle. Because we can decompose the bundle and the fiber, we can decompose \dsection\ as 
\begin{equation}
\label{eq:data_section_decompose}
\dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
\end{equation}
where each section $\dsection_i$ is a variable or set of variables. This allows for accessing the data component wise in addition to accessing the data in terms of its location over \dbase.

\begin{figure}[H]
    \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
    \caption{ Fiber (time, temperature) with an interval \dbase\ basespace. The sections $\dsection^{(1)}$ and $\dsection^{(2)}$ are constrained such that the time variable must be monotonic, which means each section is a timeseries of temperature values. They are included in the global set of sections  $\dsection^{(1)}, \dsection^{(2)} \in \Gamma(\dtotal)$}
    \label{fig:data_sections}
\end{figure}

In the example in figure~\ref{fig:data_sections}, the fiber is $(time, \, temperature)$ as described in figure~\ref{fig:data_fiber_example} and the base space is the interval \dbase. The section $\dsection^{(1)}$ resolves to a series of monotonically increasing in time records of (time, temperature) values. Section $\dsection^{(2)}$ returns a different timeseries of (time, temperature) values. Both sections are included in the global set of sections $\dsection^{(1)}, \dsection^{(2)} \in \Gamma(\dtotal)$.

>>>>

\subsubsection{Variables in Fiber Space \dfiber}
\label{sec:math:data:fiber}
To formalize the structure of the data components, we use notation introduced by Spivak \cite{spivakSIMPLICIALDATABASES} that binds the components of the fiber to variable names. This allows us to describe the components in a schema like way. Spivak constructs a set \ftotal\ that is the disjoint union of all possible objects of types $\{\ftype_0, \ldots, \ftype_m\} \in \ftypes$, where \ftypes\ are the data types of the variables in the dataset. He then defines the single variable set \fttype\ 
\begin{equation}
    \label{eq:data_types}
    \begin{tikzcd}
        \fttype \arrow[r] \arrow[d, "\pi_{\fsection}"'] & \ftotal \arrow[d, "\pi"] \\
        \fnames \arrow[r, "\fsection"']                          & \ftypes       
    \end{tikzcd}
\end{equation}
which is \ftotal\ restricted to objects of type \ftype\ bound to variable name \fname. The \fttype\ lookup is by name to specify that every component is distinct, since multiple components can have the same type \ftype. Given \fsection, the fiber for a one variable dataset is
\begin{equation}
    \dfiber = \ftotal_{\fsection(\fname)} = \ftotal_{\ftype} 
\end{equation}
where \fsection\ is the schema that binds a variable name \fname\ to its datatype \ftype. A dataset with multiple components has a fiber that is the cartesian cross product of $\ftotal_{\fsection}$ applied to all the columns:
\begin{equation}
F = \ftotal_{\fsection(\fname_{1})}\times \ldots \ftotal_{\fsection(\fname_{i})} \ldots\times \ftotal_{\fsection(\fname_{n})}
\end{equation}
which is equivalent to 
\begin{equation}
    \label{eq:math:data:fiber:decompose}
    \dfiber= \dfiber_{0} \times \ldots \times \dfiber_{i}\times\ldots\times \dfiber_{n}
\end{equation}
which allows us to decouple \dfiber\ into components $\dfiber_i$.

\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/fiber.png}
    \caption{The fiber space is the set of all combinations of all theoretically possible values of the components.  The 2D fiber \(\dfiber=\reals\times\reals\) encodes the properties of \textit{time} and \textit{temperature} components.  One dimension of the fiber encodes the range of possible values for the time component of the dataset, which is a subset of the \reals, while the other dimension encodes the range of possible values  \reals\ for the temperature component. This means the fiber is the set of points \textit{(temperature, time)}that are all the combinations of \textit{temperatature} $\times$ \textit{time}. The 3D fiber encodes points at all possible combinations of \textit{temperature}, \textit{latitude}, and \textit{longitude}.}
    \label{fig:math:data:fiber}
\end{figure}

@@

As illustrated in figure~\ref{fig:math:data:fiber}, Spivak's framework provides a consistent way to describe potentially complex components of the input data.

\subsubsection{Measurement Scales: Monoid Actions}
\label{sec:math:data:monoid}
Implementing expressive visual encodings requires formally describing the structure on the components of the fiber, which we define by the actions of a monoid on the component. In doing so, we specify the properties of the component that must be preserved in a graphic representation. A monoid \cite{Monoid2021} $\monoid$ is a set with a binary operation $\ast:\monoid \times \monoid\rightarrow \monoid$ that satisfies the axioms:
\begin{align*}
    \textbf{associativity}\; & \text{for all} a, b, c \in \monoid \, (a \bullet b) \bullet c = a \bullet (b \bullet c)\\
    \textbf{identity}\; & \text{for all } a\in \monoid,\,  e\bullet a = a 
\end{align*} 
As defined on a component of \dfiber, a left monoid action \cite{SemigroupAction2021,nlab:action} of $\monoid_i$ is a set $\dfiber_i$ with an action $\bullet: \monoid\times \dfiber_i \rightarrow \dfiber_i$ with the properties:
\begin{align*}
    \textbf{associativity}\;& \text{for all } f,g \in \monoid_i \text{ and } x\in \dfiber_i,\, f\bullet(g\bullet x) = (f\ast g) \bullet x\\
    \textbf{identity}\;& \text{for all } x\in \dfiber_i, e\in \monoid_i,\,  e\bullet x = x 
\end{align*}
The identity and associativity properties of the action denote that the action is a monoid homomorphism, which means that the group operation is preserved on both sides of the action\cite{weissteinGroupHomomorphism}.

\begin{figure}[H]
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=\textwidth]{figures/math/monoid_total.png}
        \caption{}
        \label{fig:math:data:monoid:total}
    \end{subfigure}
    \begin{subfigure}{.5\textwidth}
        \includegraphics[width=1\textwidth]{figures/math/monoid_partial.png}
        \caption{}
        \label{fig:math:data:monoid:partial}
    \end{subfigure}
    \caption{The action $\bullet$ in \autoref{fig:monoid:total} is the arrows from the partial order diagram of weather states on the left to the diagram of weather states on the right. Since the action maps the weather states to themselves, the ordering defined by the monoid $\ast$ is preserved on both sides of the action. The action in \autoref{fig:monoid:partial} is monoid homomorphism because the ordering of the weather states is the same as the ordering of the elements they are mapped to. Given $sun \geq clouds\geq rain\, lightining$ on the right, the action  $sun\, clouds \rightarrow clouds$, and $rain\, lightining \rightarrow rain$ is structure preserving because on the left $cloud \geq rain$ so the relative ordering of elements is the same as the elements they are mapped to.}
    \label{fig:math:data:monoid}
\end{figure}

One example of monoids are partial orderings on a set, such as seen in \label{fig:math:data:monoid}. Each hasse diagram of the set of weather states describes an ordering on the set; the arrow goes from the lesser value to the greater one. For example, $storm \leq rain$. In \autoref{fig:monoid:total}, the action $\bullet$ maps the elements of a set of weather states into itself by mapping them into other elements of the weather states. The action in  \autoref{fig:math:data:monoid:total}, represented as the arrows between the hasse diagrams of the weather states, maps the weather states to themselves; therefore the ordering of the weather states is identical on both sides of the action and it is therefore homomorphic. The action $\bullet$ in \autoref{fig:math:data:monoid:partial} is a monotone map\cite{fongInvitationAppliedCategory2019}
\begin{equation*}
    if\, a \leq b \; then \; \bullet(a) \leq \bullet(b) \mid a,b \in \dfiber_{i}
\end{equation*}
where the structure the action preserves is the relative, rather than exact, ordering. Since groups are monoids with invertible operations, this definition of structure is also broad enough to include the Steven's measurment scales\cite{stevensTheoryScalesMeasurement1946,leaFormalizationMeasurementScale}. Monoids are also commonly found in functional programming since the core property of monoids is composability \cite{yorgeyMonoidsThemeVariations}.

As with the fiber \dfiber\, the total monoid space \monoid\ is the cartesian product
\begin{equation}
    \label{eq:math:data:monoid:decompose}
\monoid= \monoid_{0} \times \ldots \times \monoid_{i}\times \ldots \times\ldots \monoid_{n}
\end{equation}
of each monoid $\monoid_{i}$ on $\dfiber_{i}$.  The monoid is also added to the specification of the fiber $(\fname_i,\, \ftype_i,\, \fttype\, \monoid_i)$

\subsubsection{Continuity of the Data $K$} 
\label{sec:math:data:base}
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/k_different_types.png}
    \caption{The topological base space \dbase\ encodes the continuity of the data space, for example if the data is discrete points or lies on a plane or a sphere}
    \label{fig:math:data:base:types}
\end{figure}
The base space \dbase\ acts as an indexing space, as emphasized by Butler\cite{butlerVectorBundleClassesForm1992,butlerVisualizationModelBased1989}, to express how the records in \dtotal\ are connected to each other. As shown in \autoref{fig:math:data:base:types}, 
\dbase\ can have any number of dimensions and can be continuous or discrete. The base space also does not describe anything about the dataset besides the continuity. While the base space may have components to identify the continuity, such as \textit{time, latitutde, longitude}, these labels are indexed into from \dbase\ the same as any other component. This is similar to the notion of structural \textit{keys} with associated \textit{values} proposed by Munzner\cite{munznerVisualizationAnalysisDesign2014}, but our model treats keys as a pure reference to topology. Decoupling the keys from their semantics allows the metadata to be altered; this provides for coordinate agnostic representation of the continuity and facilitates encoding of data where the independent variable may not be clear. For example the amount of snow on the ground is dependent on time of day and how much snow has fallen, and changing the coordinate system or time resolution should have no effect on how the records are connected to each other.

\begin{figure}[H]
    \includegraphics[width=\linewidth]{figures/math/k_qspace.png}
    \caption{The base space \dtotal\ is divided into fiber segments \dfiber. The base space \dbase\ acts as an index into the records in the fibers.}
    \label{fig:math:data:base:div}
\end{figure}

Formally \dbase\ is the quotient space \cite{QuotientSpaceTopology2020} of \dtotal\, meaning it is the finest space\cite{aurouxMath131Introduction} such that every $\dbasepoint \in \dbase$ has a corresponding fiber $\dfiber_k$\cite{QuotientSpaceTopology2020}. In \autoref{fig:math:data:base:div}, \dtotal\ is a rectangle divided by vertical fibers \dfiber, so the minimal \dbase\ for which there is always a mapping $\pi: \dtotal\rightarrow \dbase$ is the closed interval $\left[0,1\right]$.

As with \autoref{eq:math:data:fiber:decompose} and \autoref{eq:math:data:monoid:decompose}, we can decompose the total space into component bundles$\pi:\dtotal_i\rightarrow \dbase$ where
\begin{equation}
    \label{eq:math:data:base:decompose}
    \pi:\dtotal_1\oplus\ldots\oplus \dtotal_i \oplus\ldots \oplus \dtotal_n \rightarrow \dbase
\end{equation}
such that \(\monoid_{i}\) acts on component bundle \(\dtotal_i\). The \dbase\ remains the same because the connectivity of records does not change just because there are fewer components in each record. By encoding this continuity in the model as \dbase\, the data model now explicitly carries information about its structure such that the implicit assumptions of the visualization algorithms are now explicit.
\begin{figure}[H]
    \includegraphics[width=1\textwidth]{figures/math/base.png}
    \caption{These two datasets have the same \textit{(time, temperature)} fiber, but different continuities. The dataset on the left consists of discrete records, while the records in the dataset on the right sampled from a continuous space.} 
    \label{fig:math:data:base}
\end{figure}

The datasets in figure~\ref{fig:math:data:base} have the same fiber of (temperature, time). The dot represents a discrete base space \dbase, meaning that every dataset encoded in the fiber bundle has discrete continuity. The line is a representation of a 1D continuity, meaning that every dataset in the fiber bundle is 1D continous. By encoding this continuity in the model as \dbase\, the data model now explicitly carries information about its structure such that the implicit assumptions of the visualization algorithms are now explicit. The explicit topology is a concise way of distinguishing visualizations that appear identical, for example heatmaps and images.

\subsubsection{Data \dsection}
\label{sec:data_section}
While the projection function $\pi:\dtotal \rightarrow\dbase$ ties together the base space \dbase\ with the fiber \dfiber, a section $\dsection: \dbase\rightarrow \dtotal$ encodes a dataset. A section function takes as input location $\dbasepoint \in \dbase$ and returns a record $\delement \in \dtotal$. For example, in the special case of a table \cite{spivakSIMPLICIALDATABASES}, \dbase\ is a set of row ids, \dfiber\ is the columns, and the section \dsection\ returns the record \delement\ at a given key in \dbase. For any fiber bundle, there exists a map
\begin{equation}
    \begin{tikzcd}
        \dfiber \arrow[r, hook] & \dtotal \arrow[d, "\pi"'] \\
                          & \dbase \arrow[u, "\dsection"', bend right]
    \end{tikzcd}
\end{equation}
 such that $\pi(\dsection(\dbasepoint)) = \dbasepoint$. The set of all global sections is denoted as $\Gamma(\dtotal)$. Assuming a trivial fiber bundle $\dtotal = \dbase \times \dfiber$, the section is 
\begin{equation}
    \label{eq:section_return}
    \dsection(\dbasepoint) = (\dbasepoint, (g_{\dfiber_{0}}(\dbasepoint), \ldots, g_{\dfiber_{n}}(\dbasepoint)))
\end{equation}
where $g: \dbase \rightarrow \dfiber$ is the index function into the fiber. This formulation of the section also holds on locally trivial sections of a non-trivial fiber bundle. Because we can decompose the bundle and the fiber (\autoref{eq:math:data:base:decompose}, \autoref{eq:math:data:fiber:decompose}), we can decompose \dsection\ as 
\begin{equation}
\label{eq:math:data:section:decompose}
\dsection= (\dsection_0,\ldots, \dsection_i, \dots, \dsection_n) 
\end{equation}
where each section $\dsection_i$ maps into a record on a component $\dfiber_i \in \dfiber$. This allows for accessing the data component wise in addition to accessing the data in terms of its location over \dbase.

\begin{figure}[H]
    \includegraphics[width=1\linewidth]{figures/math/fiberbundle.png}
    \caption{ Fiber (time, temperature) with an interval \dbase\ basespace. The sections $\dsection^{(1)}$ and $\dsection^{(2)}$ are constrained such that the time variable must be monotonic, which means each section is a timeseries of temperature values. They are included in the global set of sections  $\dsection^{(1)}, \dsection^{(2)} \in \Gamma(\dtotal)$}
    \label{fig:math:data:sections}
\end{figure}

In \autoref{fig:math:data:sections}, the fiber is the same encoding of $(time, \, temperature)$ illustrated in \autoref{fig:math:data:fiber}, and the base space is the interval \dbase\ shown in \autoref{fig:math:data:base}. The section $\dsection^{(1)}$ is a function that for a point \dbasepoint\ returns a record in the fiber \dfiber. The section applied to a set of points in \dbase\ resolves to a series of monotonically increasing in time records of \textit{(time, temperature)} values. Section $\dsection^{(2)}$ returns a different timeseries of \textit{(time, temperature)} values. Both sections are included in the global set of sections $\dsection^{(1)}, \dsection^{(2)} \in \Gamma(\dtotal)$.
